---
title: "Basic Stats Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic Stats Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introducing the Basic Stats package

This packages is a collection of simple statistical functions that can be applied generally and for use in genetic data analysis. Examples include finding means and maximums, Fst generation, and normality tests.

## Intallation Instructions

```{r setup}
devtools::install_github("Hustler-1/Project_3")
library(basicstats)
library(tibble)
library(readr)
library(googlesheets4)
library(tidyverse)
library(hierfstat)
```

## Where to get the data

+ cam and crab datasets in the the data and vignettes folders
+ all other needed data is built into the functions themselves

## Using maxR

```{r warning=FALSE}
basicstats::maxr(path = "/cloud/project/data/crabs_test - crabs.csv", column = rear_width)#we can load in the data that we want to get a max for and then specify which data column we intend to get a max of.
```
## Using meanR

```{r warning=FALSE}
basicstats::meanr(path = "/cloud/project/data/crabs_test - crabs.csv", column = rear_width)#we can use the same format with meanr to get a mean value for the same data column.
```

## Using cleanR

```{r warning=FALSE}
surveys = read_csv("portal_data_joined.csv")
view(surveys)#here we can see that we have a data set with many NA values.
surveys_clean = basicstats::cleanR(data = surveys, clean_names = FALSE)#we can quickly and cleanly remove those NA throughout the whole data set.
view(surveys_clean)
#But what if we have a data set in which the naming conventions are not efficient e.g. using spaces in column names?
gs4_deauth()
cam = read_sheet("https://docs.google.com/spreadsheets/d/1esrbC-QqoufjlDHXtIYtDcsXovC_x6o6Z4JfBPKvg3s/edit?usp=sharing")
view(cam)
cam_names = basicstats::cleanR(data = cam, clean_names = TRUE)# we can use the clean_names argument in cleanR to correct those issues.
view(cam_names)#now we have replaced those pesky spaces with nice periods!
```

## Using normalitycheckR

```{r warning=FALSE}
#It is important to check the normality of your data when you plan on doing any statistical tests on data sets.
#You could run every test individually or you could use normalitycheckR to run the 3 most commonly used tests for normally distributed data.
#This function will run the Shapiro-Wilk test, Kolmogorov-Smirnov test and the Anderson-Darling test on any data you give it.
norm_test = rnorm(5000)#First lets make a test distribution using 5000 randomly generated data points. 
basicstats::normalitycheckR(norm_test)#lets see what we get.
#We can see from the output that all of the three tests came back saying that the data is normally distributed which we expected.
crabs = read_csv("crabs_test - crabs.csv")
basicstats::normalitycheckR(crabs$rear_width)#Now lets try this with real data
#Here we find that while the data was capable of passing the Shapiro-Wilk test it failed on the Kolmogorov-Smirnov test.
#this means that when compared to a normal distribution with the same mean the values in rear_width are significantly different.
```

## Using boxR

```{r warning=FALSE}
#Another way to check for normality or to just represent data is to plot it. One of the most common ways is to generate a boxplot.
basicstats::boxR(data = crabs, x = sex, y = rear_width, title = "sex vs. rear_width", xname = "sex", yname = "rear_width", points = FALSE, savename = "test.png")#Here we can simply and easily generate a basic boxplot looking at how rear width in crabs compares to the sex. Just enter "n" when prompted in the console.
#This function allows you to specify all of you axis titles as well as the title of the whole plot when generating.
#It also allows for easy saving of the plot.
basicstats::boxR(data = crabs, x = sex, y = rear_width, title = "sex vs. rear_width", xname = "sex", yname = "rear_width", points = FALSE, savename = "test.png")#If you simply enter "y" when prompted you will notice that that the plot will have been saved in you main directory.
#Another thing you might want is to see where the actuall data points are on this plot.
basicstats::boxR(data = crabs, x = sex, y = rear_width, title = "sex vs. rear_width", xname = "sex", yname = "rear_width", points = TRUE, savename = "test.png")#We can do this by changing the points argument to TRUE and you can choose whether or not to save this one.
```

## Using chunkR

```{r eval=FALSE}
#Now lets say you've ran your tests, gotten some simple statistics, and plotted some of the data. What if you wanted to manipulate your whole data set?
test = basicstats::chunkR(data = cam_names, selection = c("Common.Name", "Count", "Subproject"), group = Common.Name, sum.name = "total")#Here we can use the cam_names data set we generated earlier to create a sub-chunk of the data set that gives use the counts of every animal spotted by the cams.
view(test)#Now we can see the total counts of all observed species by their common names
#It is important to note that the object in which you are grouping in the group argument can only be one thing. This function cannot handle grouping multiple things at this time.
```

## Using FstPaiR

```{r fig.height=8, fig.width=8, warning=FALSE}
library("hierfstat")
data(gtrunchier)#first we need to get the data from hierfstat.
basicstats::FstpaiR(data = gtrunchier, pcoa = FALSE)#If you have population genetic data you can use this function to get the Weir and Nei pairwise Fst values
basicstats::FstpaiR(data = gtrunchier, pcoa = TRUE)#After getting those values you might want to plot that data to get a visualization of it. So we can use the pcoa argument to generate a pcoa plot of the data along with our raw pairwise values
```

## Using PCAPlotR

```{r eval=FALSE}
#Along with your pairwise values you might want to run a PCA analysis on the same data.
basicstats::PCAplotR(data = iris, loadings = FALSE, clusters = FALSE)#We can use PCAplotR to quickly generate a PCA plot of the data. But what if we want the loading values to appear on the plot?
basicstats::PCAplotR(data = iris, loadings = TRUE, clusters = FALSE)#Using the loadings argument we can add the loading values as well as their directions to the plot. But what about clusters?
basicstats::PCAplotR(data = iris, loadings = FALSE, clusters = TRUE)#Using the clusters argument we can add colored clusters to the plot around the data points
#It is important to note that for this function you cannot have both loadings and clusters.
```
